Efficient pedigree recording for fast population genetics simulation

### 2017-12-27 forwards_paper.tex
get rid of yet another sentence starting with This

The tree sequences produced by default by this method
are very compact, storing genotype \emph{and} genealogical information
in a small fraction of the space taken by a compressed VCF file.
The format is also highly efficient to process,
leading to advantages for downstream analysis as well.
This is because many algorithms to compute statistics of interest for population genetics
are naturally expressed in terms of tree topologies,
and so can be quickly computed from the trees underlying the tree sequence format.
For example, pairwise nucleotide diversity $\pi$, is the average density of
differences between sequences in the sample.
To compute this directly from sequence data at $m$ sites in $n$ samples
requires computing allele frequencies, taking $O(nm)$ operations.
By using the locations of the mutations on the marginal trees,
and the fact that these are correlated,
sequential tree algorithms similar to those in~\citep{kelleher2016efficient}
can do this in roughly $O(n + m \log n)$ operations.
The \msprime{} API provides a method to compute $\pi$ among arbitrary subsets of the
samples in a tree sequence, which took about 1.2 seconds
applied to the example tree sequence above with 1.1 million mutations
in 200 megabases for 500,000 samples.
Calculating $\pi$ from the numeric genotype matrix using numpy took \plr{HOW MANY} seconds.

>>>>

The tree sequences produced by default by this method
are very compact, storing genotype \emph{and} genealogical information
in a small fraction of the space taken by a compressed VCF file.
The format is also highly efficient to process,
leading to advantages for downstream analysis as well.
The dowstream advantages are because many algorithms to compute statistics of interest for population genetics
are naturally expressed in terms of tree topologies,
and so can be quickly computed from the trees underlying the tree sequence format.
For example, pairwise nucleotide diversity $\pi$, is the average density of
differences between sequences in the sample.
To compute this directly from sequence data at $m$ sites in $n$ samples
requires computing allele frequencies, taking $O(nm)$ operations.
By using the locations of the mutations on the marginal trees,
and the fact that these are correlated,
sequential tree algorithms similar to those in~\citep{kelleher2016efficient}
can do this in roughly $O(n + m \log n)$ operations.
The \msprime{} API provides a method to compute $\pi$ among arbitrary subsets of the
samples in a tree sequence, which took about 1.2 seconds
applied to the example tree sequence above with 1.1 million mutations
in 200 megabases for 500,000 samples.
Calculating $\pi$ from the numeric genotype matrix using numpy took \plr{HOW MANY} seconds.
